Smoothing seismicity techniques applied to seismic source characterization and probabilistic seismic hazard analysis in Brazil

### 2015-02-03 d01-terms.tex
minnor corrections, references and symbol translations

%-----------------------------------------
% terms
%-----------------------------------------
\newglossaryentry{equake}
{
	name={terremoto},
	description={ruptura de alguma estrutura geológica},
	plural={terremotos}
}

\newglossaryentry{seismicity}
{
	name={sismicidade},
	description={ocorrência dos tremores},
}

\newglossaryentry{hypocenter}
{
	name={hipocentro},
	description={representação geométrica do ponto no espaço, onde 
		se iniciou o \gls{rupture_process} da \gls{crust}},
	plural={hipocentros}
}

\newglossaryentry{epicenter}
{
	name={epicentro},
	description={projeção ortogonal, sobre a superfície, do \gls{hypocenter}},
	plural={epicentros}
}

\newglossaryentry{seismotectonic}
{
	name={sismotectônica},
	description={o estudo das relações entre os \glspl{equake} e a \gls{tectonic} recente de uma região.
				 Procura compreender exatamente quais são os mecanismos que levam à uma ruptura geológica 
				 e são responsáveis pela
				 \gls{seismic_activity} em uma certa área. Isso é feito analisando-se de forma combinada 
				 registros recentes de tectonismo global e regional, 
				 considerando também evidências históricas e geomorfológicas},
}

@@

\newglossaryentry{tectonic}
{
	name={tect\^onica},
	description={disciplina científica focada nos processos respons\'aveis 
				 pela cria\c{c}\~ao e transforma\c{c}\~ao das estruturas geológicas da Terra e de outros planetas.},
	plural={tect\^onicas}
}

@@

\newglossaryentry{smoothing}
{
	name={técnicas de suavização},
	description={consiste em capturar importantes feições do conjunto de dados,
				 eliminando ruídos e outras estruturas de curto comprimento de onda
				 presentes nos dados},
}

\newglossaryentry{kernel_function}
{
	name={função de núcleo},
	description={funções n-dimensionais, cuja integral em todo o domínio resulta em 1,
				 podendo ser usadas como estimativas para 
				 funções de densidade de probabilidade},
	plural={funções de kernel},
}

\newglossaryentry{seismic_rate}
{
	name={taxa de sismicidade},
	description={taxa com que terremotos são produzidos por determinada \gls{seismic_source}},
	plural={taxas de sismicidade},
}

\newglossaryentry{seismic_activity}
{
	name={atividade sísmica},
	description={frequ{ê}cia de ocorr{ê}ncia de \glspl{equake}},
}

@@

\newglossaryentry{seismic_source}
{
	name={fonte sísmica},
	description={estrutura geológica capaz de produzir tremores de terra},
	plural={fontes sísmicas}
}

\newglossaryentry{point_source}
{
	name={fonte sísmica pontual},
	description={representação geométrica por um ponto, de uma fonte sísmica},
	plural={fontes sísmicas pontuais},
}

\newglossaryentry{gmpe}
{
	name={GMPE},
	description={Equação de predição do movimento do chão},
	plural={GMPEs},
}

\newglossaryentry{area_source}
{
	name={fonte sísmica poligonal},
	description={representação geométrica por um polígono em superfície, 
				 de uma fonte sísmica},
	plural={fontes sísmicas poligonais},
}

@@

\newglossaryentry{isocista}
{
	name={isocista},
	description={curva que une valores de mesma intensidade sísmica},
}

>>>>

%-----------------------------------------
% terms
%-----------------------------------------
\newglossaryentry{equake}
{
	name={earthquake},
	description={rupture in some geological strutcture},
	plural={earthquakes}
}

\newglossaryentry{seismicity}
{
	name={seismicity},
	description={earthquake occurrence},
}

\newglossaryentry{hypocenter}
{
	name={hypocenter},
	description={geometrical representation of the starting point of the \gls{rupture_process} on the \gls{crust}},
	plural={hypocenters}
}

\newglossaryentry{epicenter}
{
	name={epicenter},
	description={orthogonal projecton over Earth surface of \gls{hypocenter}},
	plural={epicentros}
}

\newglossaryentry{seismotectonic}
{
	name={seismotectonic},
	description={o estudo das relações entre os \glspl{equake} e a \gls{tectonic} recente de uma região.
				 Procura compreender exatamente quais são os mecanismos que levam à uma ruptura geológica 
				 e são responsáveis pela
				 \gls{seismic_activity} em uma certa área. Isso é feito analisando-se de forma combinada 
				 registros recentes de tectonismo global e regional, 
				 considerando também evidências históricas e geomorfológicas},
}

@@

\newglossaryentry{tectonic}
{
	name={tectonic},
	description={disciplina científica focada nos processos respons\'aveis 
				 pela cria\c{c}\~ao e transforma\c{c}\~ao das estruturas geológicas da Terra e de outros planetas.},
	plural={tectonics}
}

@@

\newglossaryentry{smoothing}
{
	name={smoothing techniques},
	description={consiste em capturar importantes feições do conjunto de dados,
				 eliminando ruídos e outras estruturas de curto comprimento de onda
				 presentes nos dados},
}

\newglossaryentry{kernel_function}
{
	name={kernel function},
	description={n-dimentional function, which integral over whole domain is equal one and could be used as a probability density estimation.},
	plural={kernel functions},
}

\newglossaryentry{seismic_rate}
{
	name={seismic rate},
	description={rate within earthquakes are generated in some \gls{seismic_source}},
	plural={seismic rates},
}

\newglossaryentry{seismic_activity}
{
	name={seismic activity},
	description={frequency of \glspl{equake} occurence},
}

@@

\newglossaryentry{seismic_source}
{
	name={seismic source},
	description={geological structure able to produce \glspl{equake}},
	plural={seismic sources}
}

\newglossaryentry{point_source}
{
	name={point seismic source},
	description={geometrical representation as point of some seismogenic source},
	plural={point seismic sources},
}

\newglossaryentry{gmpe}
{
	name={GMPE},
	description={ground motion prediction equation},
	plural={GMPEs},
}

\newglossaryentry{area_source}
{
	name={area seismic source},
	description={representação geométrica por um polígono em superfície, 
				 de uma fonte sísmica},
	plural={farea seismic sources},
}

@@

\newglossaryentry{isocista}
{
	name={isocist},
	description={border of a region with the same seismic intensity from the same \gls{equake}},
}

### 2015-02-03 d03-acronyms.tex
minnor corrections, references and symbol translations

\newacronym{psha}{PSHA}{Análise Probabilística de Ameaça Sísmica}
\newacronym{dsha}{DSHA}{Análise Determinística de Ameaça Sísmica}
\newacronym{GMPE}{\gls{gmpe}}{\glsdesc{gmpe}}
\newacronym{mfd}{MFD}{Distribuição de Frequência e Magnitude}
\newacronym{gr}{GR}{Gutenberg-Richter}
\newacronym{cnn}{CNN}{Acoplamento dos Vizinhos mais Próximos}
\newacronym{va}{v.a.}{variável aleatória}
\newacronym{iid}{i.i.d.}{independentes e identicamente distribuidas}
\newacronym{pdf}{densidade}{função de densidade de probabilidade}
\newacronym{pmf}{pmf}{função de distribuição acumulada de probabilidade}

\newacronym{pga}{PGA}{máxima aceleração do chão}

@@

\newacronym{hl}{HazardLib}{Biblioteca de Ameaça}
\newacronym{rl}{RiskLib}{Biblioteca de Risco}
\newacronym{nrml}{NRML}{Natural-hazard and Risk Markup Language}
\newacronym{hmtk}{HMTK}{Hazard Modeller's Toolkit}
\newacronym{oqp}{oq-platform}{Plataforma web de Interação com o OQ}
\newacronym{oqe}{oq-engine}{Motor de Cálculo do OQ}

\newacronym{obsis}{ObSis}{Observatório Sismológico}
\newacronym{unb}{UnB}{Universidade de Brasilia}
\newacronym{iag}{IAG}{Instituto de Astronomia, Geofísica e Ciências Atmosféricas}
\newacronym{usp}{USP}{Universidade de São Paulo}
\newacronym{ipt}{IPT}{Instituto de Pesquisas Tecnológicas}
\newacronym{unesp}{UNESP}{Universidade Estatual Paulista}
\newacronym{ufrn}{UFRN}{Universidade Federal do Rio Grande do Norte}

\newacronym{bsb}{BSB}{Boletim Sísmico Brasileiro}
\newacronym{csv}{CSV}{Valores Separados por Vírgulas}

\newacronym{bsb2013}{{BSB-2013.08}}{Brazilian Seismic Bulletin version 2013.08}
\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{Catálogo ISC-GEM para a América do
Sul}

>>>>

\newacronym{psha}{PSHA}{Probabilistic Seismic Hazard Analysis}
\newacronym{dsha}{DSHA}{Deterministic Seismic Hazard Analysis}
\newacronym{GMPE}{\gls{gmpe}}{\glsdesc{gmpe}}
\newacronym{mfd}{MFD}{Magnitude Frequency Distribution}
\newacronym{gr}{GR}{Gutenberg-Richter}
\newacronym{cnn}{CNN}{Coupled-Nearest-Neighbour}
\newacronym{va}{r.v.}{random variable}
\newacronym{iid}{i.i.d.}{independent and identicaly distributed}
\newacronym{pdf}{density}{probability density function}
\newacronym{pmf}{pmf}{probability mass function}
\newacronym{kde}{KDE}{Kernel Density Estimation}

\newacronym{pga}{PGA}{peak ground-motion acceleration}

@@

\newacronym{hl}{HazardLib}{hazard library}
\newacronym{rl}{RiskLib}{risk library}
\newacronym{nrml}{NRML}{Natural-hazard and Risk Markup Language}
\newacronym{hmtk}{HMTK}{Hazard Modeller's Toolkit}
\newacronym{oqp}{oq-platform}{Plataforma web de Interação com o OQ}
\newacronym{oqe}{oq-engine}{OQ hazard and risk calculator}

\newacronym{obsis}{ObSis}{Seismological Observatory}
\newacronym{unb}{UnB}{Brasilia University}
\newacronym{iag}{IAG}{Institute of Astronomy, Geophysics and Atmosferical Sciences}
\newacronym{usp}{USP}{São Paulo University}
\newacronym{ipt}{IPT}{Institute of Technological Research}
\newacronym{unesp}{UNESP}{São Paulo State University}
\newacronym{ufrn}{UFRN}{Federal University of Rio Grande do Norte}

\newacronym{bsb}{BSB}{Brazilian Seismic Bulletin}
\newacronym{csv}{CSV}{Common separated values}

\newacronym{bsb2013}{{BSB-2013.08}}{Brazilian Seismic Bulletin version 2013.08}
\newacronym{bsb2014}{{BSB-2014.11}}{Brazilian Seismic Bulletin version 2011.11}
\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{ISC-GEM catalog for South America}

### 2015-02-03 manuscript.tex
minnor corrections, references and symbol translations

% Author names in capital letters:
\authorrunninghead{PIRCHINER M.}

@@

Three smoothing seismicity techniques are reviewed and their application tothe Brazilian seismicity and a seismic hazard assesment are discussed.

All these smoothing methods are based on a kernel density estimation. The main difference between them is the bandwidth selection process.

The method of \citet{frankel_1995} uses a \emph{fixed} bandwidth in space. \citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}. \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.
This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
(not indiscriminately REFERENCE) as a grid of points seismic sources.

@@

The Brazilian seismic hazard building code (REFERENCE) [ABNT] was based on the GSHAP results for South America, which is the last published model for the whole region. Howerver, some significant seismic areas (Porto dos Gaúchos [REFERENCIA]), are not represented in the GPSHAP-based building codes.

A significant number of earthquakes has been recorded in the last decades allowing a better definition of seismic zones. For this reason an updated hazard map is needed.

@@

Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on kernel density estimation which results in a grid of smoothed seismicity.

This kernel density estimation or the smoothed seismicity grid allow to define a grid of seismic point sources. The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.

In the kernel density estimation both its shape and its bandwidth need to be defined. Most kernels are gaussian or power-laws in one of their variants. The main distinction between the methods described here is the bandwidth selection process.

@@

For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation of Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT). From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica. Since 1995, the Bulletin is maintained mainly by University of São Paulo under the same cooperation network in addition to Universidade Estadual Paulista (UNESP). Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.

@@

Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude. For historical events magnitudes were estimated from macroseismic data (felt area and maximum intensity) \citep{berrocal_1984}. Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}. REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS.. (Johnston)
% Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.

\subsection{Catalog check}

Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour. Most events have no depth determination and the well determined depths (from pP or local networks) are mostly less than 10km.

From the figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night. This could suggest some influence of daily noise level.

The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara [REFERENCE, TAKEYA et al 1989] and Porto dos Gaúchos [REFERENCE, BARROS 2009]. The increasing number of records after 1960 is due to the operation of global, regional and local stations.

@@

To preserve the Poisson assumption about the independence of the events several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a. The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uental \citep{marsan_david_2012}
(REVER REFERENCIA DO PROPRIO gruental ou CORRIGIN POIS NAO EH MARSAN DAVID mas Thomas van Stiphout, Jiancang Zhuang, and David Marsan].
In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uental distance window.

@@

Figure \ref{fig_occurrence} presents the magnitude and frequency distribution.
A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.

@@

- completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
- se conhecia metodos para lidar com completeza espacial
- também se conhece o método multiescala (provavel bom solucionador)
- método de Stepp
- mas usa-se uma completeness temporal simples.

@@

Arthur \citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$. $n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.

@@

The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection. These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$. These weights could easily consider space and time completenes and \emph{b}-value fluctuations. The expression used to compute it:
\begin{equation}
	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
	\right] } },
	\label{eq:helms_wi}
\end{equation}
where \gls{sym:wi} is the \glsdesc{sym:wi} at location $\boldsymbol{r}$ on instant $t$,
	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
	  \gls{sym:Mc_rt} is the \glsdesc{sym:Mc_rt},
	  \gls{sym:Md} is the \glsdesc{sym:Md}.

@@

The method implemented to compute the space and time bandwidths for each earthquake was the Couppled-Nearest-Neighbour \citep{helmstetter_2012}:
\begin{equation}
	\ensuremath{
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
	 \gls{sym:dk} is the \glsdesc{sym:dk} e
	 \gls{sym:hk} is the \glsdesc{sym:hk}.

The bandwidths are defined locally. Could be small in high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.

\subsubsection{Stationary seismic rate}
After compute the model parameters and completely define it, the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ have to be computed.
The way to do it proposed by \citet{helmstetter_2012} was to get the median
on $\boldsymbol{r}_0$ over all considered time window:
\begin{equation}
	\ensuremath{
		\bar{R}(\boldsymbol{r}_0) = \text{Median}\left[R(\boldsymbol{r}_0, t)\right].
	}
	\label{eq:stationary_rate}
\end{equation}

The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures.

@@

The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}. To optimize these parameters the catalog is divided in two parts: one for learning about the parameters and other to test them.

If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the consider time frame is
\begin{equation}
	\ensuremath{
		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
	}
	\label{eq:poisson_probability}
\end{equation}

Then, over all cells, the (log) likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed on testing catalog is written as
\begin{equation}
	\ensuremath{
		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
	}
	\label{eq:loglik}
\end{equation}
where \gls{sym:Np} is the \glsdesc{sym:Np} and
	  \gls{sym:nxy} is the \glsdesc{sym:nxy}.

@@

%------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%------------------------------------------------------------------------------
\begin{acknowledgments}
The author thanks specially to Marcelo Assumpção for the support and critical discussion as well all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
\end{acknowledgments}

@@

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
		\\
		\footnotesize(b)
		& \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2013_pp_decluster}
		\end{tabular}
	\end{table}
	\caption{Decluster evaluation (a) and the clusters map (b).}
	\label{fig_decluster}
	\end{center}
\end{figure}

\begin{figure}
	\includegraphics[width=1.0\linewidth]{z_img_hmtk_bsb2014_11_completeness}
	\caption{Annual earthquake recurrence.}
	\label{fig_occurrence}
\end{figure}

@@

>>>>

% Author names in capital letters:
\authorrunninghead{PIRCHINER M. and ASSUMP\c{C}\~AO M.}

@@

Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assesment are discussed.

All these smoothing methods are based on a \gls{kde}. The main difference between them is the bandwidth selection process.

The method of \citet{frankel_1995} uses a \emph{fixed} bandwidth in space. \citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}. \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.

This last one was proposed initialy to take care of the background seismicity rate in the context of long-term earthquake forecast and the time-independent assumption about the annual background seismic rate could be used
(not indiscriminately REFERENCE) as a grid of points seismic sources.

@@

The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{shedlock_tanner_1999}, which is the last published model for the whole region. Howerver, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GPSHAP-based building codes.

A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better definition of seismic zones. For this reason an updated hazard map is needed.

@@

Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology could be used to assess seismic hazard using expert pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.

The seismicity \gls{kde} or the smoothed seismicity grid allow to define a grid of seismic point sources. The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.

To perform the \gls{kde}, both its shape and bandwidth needs to be defined. Most kernels are gaussian or power-laws or one of their variants. The main distinction between the methods described here is the bandwidth selection process.

@@

For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation of Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT). From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica. Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP). Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.

@@

Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude. For historical events magnitudes were estimated from macroseismic data (felt area and maximum intensity) \citep{berrocal_1984}. Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}. REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS. (Johnston)
% Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.

\subsection{Catalog checking}

Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour. Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10km.

Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night. This could suggest some influence of daily noise level.

The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara \citep{takeya_1989} and Porto dos Gaúchos \citep{barros_2009}. The increasing number of records after 1960 is due to the operation of global, regional and local stations.

@@

To preserve the Poissonian assumption about the independence of the events, several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a. The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uenthal \citep{van_stiphout_2012}.
In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window.

@@

Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.

@@

\begin{itemize}
	\item completeza é problema temporal mas nesse caso tambem espacial nao negligenciavel
	\item se conhecia metodos para lidar com completeza espacial
	\item também se conhece o método multiescala (provavel bom solucionador)
	\item conhece-se o método de Stepp
	\item mas usa-se uma completeness temporal simples
\end{itemize}

@@

\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a "correlation distance" $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a gaussian kernel:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$. $n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.

@@

The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection. This is the expression used to compute it:
\begin{equation}
	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
	\right] } },
	\label{eq:helms_wi}
\end{equation}
where \gls{sym:wi} is the \glsdesc{sym:wi} factor computed on location $\boldsymbol{r}$ on instant $t$,
	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
	  \gls{sym:Mc_rt} is the \glsdesc{sym:Mc_rt},
	  \gls{sym:Md} is the \glsdesc{sym:Md}.

These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$. They also could easily take into account the space and time completenes and \emph{b}-value fluctuations.

@@

The method implemented to compute the space and time bandwidths for each earthquake was the \gls{cnn} \citep{helmstetter_2012} expressed as:
\begin{equation}
	\ensuremath{
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
	 \gls{sym:dk} is the \glsdesc{sym:dk} and
	 \gls{sym:hk} is the \glsdesc{sym:hk}.

The bandwidths are defined locally by this simple optimization process. It could be small on high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.

\subsubsection{Stationary seismic rate}
After compute the model parameters it is completely defined, and the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ can be computed on the way proposed by \citet{helmstetter_2012} taking the median value for some location $\boldsymbol{r}_0$ over all considered time window:
\begin{equation}
	\ensuremath{
		\bar{R}(\boldsymbol{r}_0) = \text{Median}\left[R(\boldsymbol{r}_0, t)\right].
	}
	\label{eq:stationary_rate}
\end{equation}

The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures. This is one of the most important achievement of this methodology.

@@

The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}. To optimize these parameters the catalog is divided in two parts: one for learning about the parameters and other to test them. The target catalog will able us to check the model prediction performance derived from its chosen parameters. The best model parameters will maximize the prediction model capacity likely the target catalog.

If the earthquake occurrence could be modelled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
\begin{equation}
	\ensuremath{
		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
	}
	\label{eq:poisson_probability}
\end{equation}

Then, over all cells, the log-likelihood, to be maximized, between model (from the learning catalog) prediction and earthquakes observed (on testing catalog) is written as
\begin{equation}
	\ensuremath{
		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
	}
	\label{eq:loglik}
\end{equation}
where \gls{sym:Np} is the \glsdesc{sym:Np} and
	  \gls{sym:nxy} is the \glsdesc{sym:nxy}.

@@

%------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%------------------------------------------------------------------------------
\begin{acknowledgments}
The author thanks specially to 
%Marcelo Assumpção for the support and critical discussion as well 
all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
\end{acknowledgments}

@@

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.45\textwidth]{z_img_hmtk_bsb2014_11_decluster}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.40\textwidth]{z_img_hmtk_bsb2014_11_clusters}
		\end{tabular}
	\end{table}
	\caption{Decluster evaluation (a) and the clusters map (b).}
	\label{fig_decluster}
	\end{center}
\end{figure}

\begin{figure}
    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_completeness}
    \caption{Completeness}
    \label{fig_completeness}
\end{figure}

\begin{figure}
    \includegraphics[width=.45\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
    \caption{Annual earthquake recurrence.}
    \label{fig_occurrence}
\end{figure}

@@

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.45\textwidth]{z_img_hi_histogram}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.45\textwidth]{z_img_di_histogram}
		\end{tabular}
	\end{table}
	\caption{(a) $h_i$ and (b) $d_i$ distributions from \gls{bsb}.}
	\label{fig_hidi_histogram}
	\end{center}
\end{figure}
